### 2017-01-16 2017mergingmerges/section/01-Introduction.tex
intro prose

\begin{code}
uniquesUnion (file1 file2 : Stream Id) : (Stream Id, Stream Id)
 = let uniques = group file1
       union   = group (merge file1 file2)
   in (uniques, union)
\end{code}

@@

>>>>

\begin{code}
uniquesUnion (file1 file2 : Stream Id) : (Stream Id, Stream Id)
 = let uniques = group file1
       merged  = merge file1 file
       union   = group merged
   in (uniques, union)
\end{code}

@@

Sadly, stream fusion~\cite{coutts2007stream} and other forms of shortcut fusion~\cite{jones2001playing} cannot fuse this example because there are multiple consumers.
Synchronous dataflow languages~\cite{mandel2010lucy} and data flow inspired fusion~\cite{lippmeier2013data} cannot fuse this because they do not support value-dependent input patterns.
Polarised data flow fusion~\cite{lippmeier2016polarized} is able to fuse this case, but the program must be manually rewritten by categorising streams into pull or push and adding extra duplicate combinators.

Our approach is to treat each combinator as a separate process with communication channels between them.
Each process is a simple sequential imperative program with commands such as @pull@ for reading from an input channel and @push@ for writing an output.
Multiple processes form a network where processes are executed concurrently.
Fusion, then, involves taking a network of concurrent processes and choosing a particular evaluation order, condensing them to a single sequential process.

The process for @uniques@ is defined by the @group@ combinator, and operates thus: first it reads from its input stream @file1@ and stores that in a local variable.
It also keeps track of the last pulled value, and compares the just read value.
If they are different, it pushes the new value to its output stream @uniques@ and updates the last pulled value.
In either case, it loops back indefinitely to pull from @file1@ again.

Fusion is then rather like executing two processes concurrently by taking the (synchronised) product of two state machines: either the both machines can take the same step (both pulling from @file1@, or one pushing to @merged@ while the other pulls), or one of the machines can take an independent step which isn't in the alphabet of the other while the other `rests'.
Deadlock can occur with synchronous product when each machine is waiting on the other, and neither can progress.
We solve this by also storing in the state label information about which input channels are known to have values.
This way, rather than requiring both machines to take the same step at the same time, executing a @push@ from one machine fills the buffer of the other, allowing the other machine to read the value from the buffer at its leisure.
In order to push, the other machine must have an empty buffer.
In this way we have extended synchronised product to allow more cases without deadlock.

### 2017-01-16 2017mergingmerges/section/XX-Related.tex
intro prose

\subsection{The problem with synchronised product}
Related work on Network Fusion~\cite{fradet2004network} allows fusion of Kahn Process Networks using synchronised product to fuse pairs of processes together.
Synchronised product is: both machines can take independent steps so long as it is not in the alphabet of the other, but have to agree on any shared actions.
So if both machines pull from @xs@, the @pull xs@ message can only be executed when both machines agree.

>>>>

\subsection{The problem with synchronised product}
\label{s:Synchro}
Related work on Network Fusion~\cite{fradet2004network} allows fusion of Kahn Process Networks using synchronised product to fuse pairs of processes together.
Synchronised product is: both machines can take independent steps so long as it is not in the alphabet of the other, but have to agree on any shared actions.
So if both machines pull from @xs@, the @pull xs@ message can only be executed when both machines agree.
